---
title: "SpeedTest Results"
output: html_notebook
---

This notebook gets the latest results for SpeedTest results that have been run on a recurring schedule and logged to a Google Sheet and then plots them. The source code for both the logging script, this output, and a Shiny app for output is available [on Github](https://github.com/gilliganondata/speedtest-log).

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

library(tidyverse)
library(lubridate)
library(scales)
library(googlesheets4)

# Get the ID for the Google Sheet where the results are stored. The technique below expects this
# ID to be recorded in .Renviron more another environment file in a variable called "SPEEDTEST_GSHEET."
# Alternatively, you can just hardcode the Google Sheets ID in a variable:
# gsheet <- "[sheet ID]"
gsheet <- Sys.getenv("SPEEDTEST_GSHEET")

# Set the timezone for displaying the results
timezone <- "America/New_York"

# Assumes the Google Sheet with the data has been made public (view only / with link)
sheets_deauth()
```

## Get the Data

The data gets read from the Google Sheets file specified in the setup. There is one dataset (sheet) for the download tests and one dataset (sheet) for the upload tests.

```{r get_data, echo = FALSE, warning = FALSE, message = FALSE}

download_data <- sheets_read(gsheet, "Download Data")
upload_data <- sheets_read(gsheet, "Upload Data")

```

## Visualize the Data

Inspect the data as two simple time-series plots:

```{r visualize, echo=FALSE, warning=FALSE, message=FALSE}
# Set up theme
theme_ts <- theme_minimal() +
  theme(plot.title.position = "plot",
        plot.subtitle = element_text(face = "italic", colour = "gray40"),
        panel.grid.major.x = element_line(),
        panel.grid.minor.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_text(colour = "gray40", margin = margin(0, 5, 0, 0, "pt")),
        axis.text.y = element_text(colour = "gray30", face = "bold", size = 10))

# Set up theme for the histogram based on the time-series theme
theme_hist <- theme_ts +
  theme(axis.title.x = element_text(colour = "gray40", margin = margin(5, 0, 0, 0, "pt")),
        axis.text.x = element_text(colour = "gray30", face = "bold", size = 10),
        axis.text.y = element_text(colour = "gray30", face = "plain", size = 10))

# Function to build a plot summary of either a download or an upload test
get_plot_ts <- function(test_data, test_type){
  
  # Update the timezone to display in the timezome specified in the setup
  test_data <- test_data %>% 
    mutate(test_time = with_tz(test_time, timezone))
  
  # Get summary statistics for each round of tests
  data_viz_df <- test_data %>% 
    group_by(test_time) %>% 
    summarize(bw_median = median(bw),
              bw_mean = mean(bw),  # Note used, but adding for giggles if we wanted to chat mean vs median
              bw_min = min(bw),
              bw_first_quartile = quantile(bw, .25),
              bw_third_quartile = quantile(bw, .75),
              bw_max = max(bw))
  
  # Get the overall median of all tests for a horizonal line
  bw_median_overall <- median(test_data$bw)
  
  # Build the plot
  gg <- ggplot(data_viz_df, 
               aes(x = test_time, y = bw_median, 
                   ymin = bw_first_quartile, ymax = bw_third_quartile)) +
    # 2nd and 3rd quartile
    geom_ribbon(fill = "gray70", alpha = 0.4) +
    # 1st and 4th quartiles
    geom_ribbon(aes(ymin = bw_min, ymax = bw_max), fill = "gray80", alpha = 0.3) +
    # Overall Median
    geom_hline(mapping = aes(yintercept = bw_median_overall),
               linetype = "dashed", colour = "gray60", size = 0.7) +
    # Median by Hour
    geom_line(color = "#BF5700", size = 1) +    # Hook 'em Horns!
    scale_x_datetime(date_breaks = "1 days", date_labels = "%m/%d") +
    labs(title = paste(test_type, "Speed (Mbps) by Hour: The Overall Median Was", 
                       round(bw_median_overall,1), "Mbps"),
         subtitle = "The Gray Bands Represent the Quartiles for the Times Recorded in Each Hourly Test",
         y = paste("Median", test_type, "Speed (Mbps)")) +
    theme_ts
  
  gg
}

gg_download_ts <- get_plot_ts(download_data, "Download")
gg_upload_ts <- get_plot_ts(upload_data, "Upload")

gg_download_ts
gg_upload_ts

```

View the data as a histogram:

```{r visualize_hist, echo=FALSE, warning=FALSE, message=FALSE}

# Set up theme for the histogram based on the time-series theme
theme_hist <- theme_ts +
  theme(axis.title.x = element_text(colour = "gray40", margin = margin(5, 0, 0, 0, "pt")),
        axis.text.x = element_text(colour = "gray30", face = "bold", size = 10),
        axis.text.y = element_text(colour = "gray30", face = "plain", size = 10))

# Function to build a histogram of either a download or an upload test
get_plot_hist <- function(test_data, test_type){
  
  # Get the overall mean of all tests a vergical line
  bw_mean <- median(test_data$bw)
  
  
  # Build the histogram
  gg <- ggplot(test_data, aes(x = bw)) +
    
    # The histogram itself
    geom_histogram(fill = "#BF5700", alpha = 0.7) +
    
    # Overall mean
    geom_vline(mapping = aes(xintercept = bw_mean),
               linetype = "dashed", colour = "gray60", size = 0.7) +
    scale_y_continuous(label = comma) +
    labs(title = paste("The Mean", test_type, "Speed Was", 
                       round(bw_mean, 1), "Mbps"),
         x = paste("Mean", test_type, "Speed (Mbps)"),
         y = "# of Tests") +
    theme_hist
  
  gg
  
}

gg_download_hist <- get_plot_hist(download_data, "Download")
gg_upload_hist <- get_plot_hist(upload_data, "Upload")

gg_download_hist
gg_upload_hist

```